Phenotype and risk assesment of insect venom anaphylaxis: a case - control study of the European Anaphylaxis Registry

### 2019-10-10 analysis/paper/paper.Rmd
changed the Fig. 2 and results secion

^1^ Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Humboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,

@@

Hypersensitivity to insect venom is a systemic reaction (anaphylaxis) in up to 0.3--7.5% of the adult population [@Bil2008]. Venom induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong systemic immunotherapy [@Sturm2017].

The need for a more precise description of the diagnosis, identification of biomarkers, and phenotypes of anaphylaxis is discussed [@JimenezRodriguez2018]. Nevertheless, in order to facilitate a precision-medicine approach [@Muraro2017] for the diagnosis of anaphylaxis --- a better understanding of its clinical subtypes is required.

Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and their infrequent and random occurence. Therefore registries, gathering clinical data from patients with a well documented (recent) history of anaphylaxis are a crucial tool in investigating this entity.

This study aimed to identify clinical patterns of venom induced anaphylaxis (VIA) regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-VIA).

@@

Due to a large number of documented reactions in the European Anaphylaxis registry - we were able to match the VIA cases to selected non-VIA cases according to sex and age to reduce the comparison bias. Subsequently, we compared the management in both groups and matched the control group according to the severity of a reaction. Propensity score matching has been performed using the "MatchIt" package for R [@matchit]. The results of the propensity score matching are in Fig. \@ref(fig:flow)B-D and supplementary Fig \@ref(fig:matchingresult).

@@

When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B).

@@

VIA displayed a specific symptom pattern. Patients, who underwent VIA, more often experienced cardiovascular symptoms (dizziness, reduced alertness, unconsciousness) than patients with other elicitors inducing anaphylaxis and less often had skin or gastrointestinal symptoms (Fig \@ref(fig:figsymptoms)A). The difference in severe hypotension frequency was especially prominent in the younger age group (under 22, Fig. S\@ref(fig:supplement)).

The pattern of organ involvement showed similarities in gastrologic, skin, and respiratory symptoms, and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml). Although, patients undergoing VIA  less often had concomitant atopic diseases (Fig \@ref(fig:figsymptoms)B).

Severe reactions of VIA were more prevalent in older patients in comparison to patients below 22 years, and in VIA cases vs. other elicitors (Fig. S\@ref(fig:sev)D). There were no differences in severity of reactions elicited by yellow-jackets and other insect species.

@@

>>>>

^1^ Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Hmuboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,

@@

@@

Due to a large number of documented reactions in the European Anaphylaxis registry - we were able to match the VIA cases to selected non-VIA cases according to sex and age to reduce the comparison bias. 
When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B). Subsequently, we compared the management in both groups and matched the control group according to the severity of a reaction. Propensity score matching has been performed using the "MatchIt" package for R [@matchit]. The results of the propensity score matching are in Fig. \@ref(fig:flow)B-D and supplementary Fig \@ref(fig:matchingresult).

@@

@@

VIA displayed a specific symptom pattern. Patients, who underwent VIA, more often experienced cardiovascular symptoms (dizziness, hypotension, unconsciousness, reduced alertness) than patients with anaphylaxis due to other elicitors and less often presented with expiratory distress, rhinitis or diarrhoea (Fig. \@ref(fig:figsymptoms)A).

Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrologic, skin, and respiratory systems, VIA more frequently involved over three organ systems (`r compare(c(tempdf[3:4,">3 organ\nsystems"]))`,<!-- , and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml). Although, patients undergoing VIA  less often had concomitant atopic diseases  -->
Fig. \@ref(fig:figsymptoms)B).

@@

## Severity 
Severe reactions of VIA were more prevalent in older patients in comparison to patients below 22 years, and in VIA cases vs. other elicitors (Fig. S\@ref(fig:sev)D). There were no differences in severity of reactions elicited by yellow-jackets and other insect species.

### 2019-10-10 analysis/paper/paper_backupV2.5.Rmd
changed the Fig. 2 and results secion

@@

Due to a large number of documented reactions in the European Anaphylaxis registry - we were able to match the VIA cases to selected non-VIA cases according to sex and age to reduce the comparison bias. 
When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B). Subsequently, we compared the management in both groups and matched the control group according to the severity of a reaction. Propensity score matching has been performed using the "MatchIt" package for R [@matchit]. The results of the propensity score matching are in Fig. \@ref(fig:flow)B-D and supplementary Fig \@ref(fig:matchingresult).

@@

@@

VIA displayed a specific symptom pattern. Patients, who underwent VIA, more often experienced cardiovascular symptoms (dizziness, hypotension, unconsciousness, reduced alertness) than patients with anaphylaxis due to other elicitors and less often presented with expiratory distress, rhinitis or diarrhoea (Fig. \@ref(fig:figsymptoms)A).

Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrologic, skin, and respiratory systems, VIA more frequently involved over three organ systems (`r compare(c(tempdf[3:4,">3 organ\nsystems"]))`,<!-- , and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml). Although, patients undergoing VIA  less often had concomitant atopic diseases  -->
Fig. \@ref(fig:figsymptoms)B).

@@

## Severity 
Severe reactions of VIA were more prevalent in older patients in comparison to patients below 22 years, and in VIA cases vs. other elicitors (Fig. S\@ref(fig:sev)D). There were no differences in severity of reactions elicited by yellow-jackets and other insect species.

>>>>

Hypersensitivity to insect venom is a systemic reaction (anaphylaxis) in up to 0.3--7.5% of the adult population [@Bil2008]. Venom induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong systemic immunotherapy [@Sturm2017].

The need for a more precise description of the diagnosis, identification of biomarkers, and phenotypes of anaphylaxis is discussed [@JimenezRodriguez2018]. Nevertheless, in order to facilitate a precision-medicine approach [@Muraro2017] for the diagnosis of anaphylaxis --- a better understanding of its clinical subtypes is required.

Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and their infrequent and random occurence. Therefore registries, gathering clinical data from patients with a well documented (recent) history of anaphylaxis are a crucial tool in investigating this entity.

This study aimed to identify clinical patterns of venom induced anaphylaxis (VIA) regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-VIA).

@@

Due to a large number of documented reactions in the European Anaphylaxis registry - we were able to match the VIA cases to selected non-VIA cases according to sex and age to reduce the comparison bias. Subsequently, we compared the management in both groups and matched the control group according to the severity of a reaction. Propensity score matching has been performed using the "MatchIt" package for R [@matchit]. The results of the propensity score matching are in Fig. \@ref(fig:flow)B-D and supplementary Fig \@ref(fig:matchingresult).

@@

When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig \@ref(fig:flow)B).

@@

VIA displayed a specific symptom pattern. Patients, who underwent VIA, more often experienced cardiovascular symptoms (dizziness, reduced alertness, unconsciousness) than patients with other elicitors inducing anaphylaxis and less often had skin or gastrointestinal symptoms (Fig \@ref(fig:figsymptoms)A). The difference in severe hypotension frequency was especially prominent in the younger age group (under 22, Fig. S\@ref(fig:supplement)).

The pattern of organ involvement showed similarities in gastrologic, skin, and respiratory symptoms, and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml). Although, patients undergoing VIA  less often had concomitant atopic diseases (Fig \@ref(fig:figsymptoms)B).

Severe reactions of VIA were more prevalent in older patients in comparison to patients below 22 years, and in VIA cases vs. other elicitors (Fig. S\@ref(fig:sev)D). There were no differences in severity of reactions elicited by yellow-jackets and other insect species.

@@
